### 2016-07-18 misc/overallText.tex
add of aux file

Database consists of 37467423 triples, 3088219 edges yield by interactions or relations, 382735 participants and 229008584 characters. Among all snapshots, 63 are ego snapshots, 53 are group snapshots; 45 have interaction edges, 89 have friendship edges; 42 have text content from messages.