### 2016-07-28 tables/nsnapshots.tex

\begin{table*}[h!]
\begin{center}
\caption{Number of snapshots from each provenance. Every snapshot is a \texttt{po:Snapshot}; there are three types of the \texttt{po:AASnapshot} class.}
\begin{tabular}{| l | c |}\hline
\textbf{snapshot provenance} & \textbf{number of snapshots} \\\hline\hline\hline
http://purl.org/socialparticipation/po/Snapshot & 117 \\\hline
\end{tabular}\end{center}
\end{table*}