### 2016-07-28 tables/nsnapshots.tex
finished rendering tables

\begin{table*}[h!]
\begin{center}
\caption{Number of snapshots from each provenance.}
\begin{tabular}{| l | c |}\hline
\textbf{social protocol} & \textbf{number of snapshots} \\\hline\hline
Algorithmic Autoregulation & 3 \\\hline
Cidade Democrática & 1 \\\hline
Email & 4 \\\hline
Facebook & 88 \\\hline
IRC & 4 \\\hline
ParticipaBR & 1 \\\hline
Twitter & 16 \\\hline\hline
all & 117 \\\hline
\end{tabular}\end{center}
\end{table*}