### 2017-01-10 tables/facebookReferences.tex
adicionada tabela de referencias do fb

\begin{table*}[h!]
\begin{center}\scriptsize
\caption{All the Facebook snapshots are either the result of individuals who downloaded
their data (and donated to the first author) or data downloaded from groups.
In the first case, it is senseless to present references. In the second
case, we present the group name and a link to a post in the group where
data and figures were delivered back to the group.}
\begin{tabular}{| l | p{9.9cm} |}\hline
\textbf{group name} & \textbf{url(s)} \\\hline
Adorno Nao Eh Enfeite & \url{https://www.facebook.com/groups/265217103529531/permalink/525654127485826/} \\\hline
Ativistas Da Inclusao Digital & \url{https://www.facebook.com/groups/423602557691243/permalink/525201037531394/} \\\hline
Ciencias Com Fronteiras & \url{https://www.facebook.com/groups/contraaexclusao/permalink/269103356558439/} \\\hline
Computer Art & \url{https://www.facebook.com/groups/computerart/permalink/259389137529870/} \\\hline
Coolmeia &  \url{https://www.facebook.com/groups/coolmeia/permalink/380091142098291/} , \url{https://www.facebook.com/groups/coolmeia/permalink/489757754464962/} \\\hline
Democracia Direta Ja & \url{https://www.facebook.com/groups/ddjbrasil/permalink/347023325397298/} \\\hline
Democracia Pura & \url{https://www.facebook.com/groups/democraciapura/permalink/310907215704321/} \\\hline
Economia & \url{https://www.facebook.com/groups/economa1/permalink/586007714743535/} \\\hline
Economia Criativa Digital & \url{https://www.facebook.com/groups/economiacriativadigital/permalink/438313682916103/} \\\hline
Educacoes E Aprendizagens XXI & \url{https://www.facebook.com/groups/geaxxi/permalink/433229973421182/} \\\hline
Latesfip & \url{https://www.facebook.com/groups/183557128478424/permalink/266610616839741/} \\\hline
Living Bridges Planet & \url{https://www.facebook.com/groups/livingbridgesplanet/permalink/352950408144951/} \\\hline
Mobilizacoes Culturais Interior SP & \url{https://www.facebook.com/groups/131639147005593/permalink/144204529082388/} \\\hline
Partido Pirata & \url{https://www.facebook.com/groups/partidopiratabrasil/permalink/10151409024509317/} \\\hline
Politicas Culturas Brasileiras & \url{https://www.facebook.com/groups/pcult/permalink/519626544747423/} \\\hline
Praca Popular & \url{https://www.facebook.com/groups/215924991863921/permalink/319279541528465/} \\\hline
Rede Tranzmidias & \url{https://www.facebook.com/groups/318333384951196/permalink/346658712118663/} \\\hline
Silicon Valley Global Network & \url{https://www.facebook.com/groups/109971182359978/permalink/589326757757749/} \\\hline
Solidarity Economy & \url{https://www.facebook.com/groups/9149038282/permalink/10151461945623283/} \\\hline
Study Group SNA & \url{https://www.facebook.com/groups/140630009439814/permalink/151470598355755/} \\\hline
Tecnoxamanismo &  \url{https://www.facebook.com/groups/505090906188661/permalink/733144993383250/} , \url{https://www.facebook.com/groups/505090906188661/permalink/733157380048678/} \\\hline
\end{tabular}\end{center}
\end{table*}