### 2019-06-30 misc/overallText.tex
adiantada adaptacao p revista data

The database consists of 34,120,026 triples, 3,172,927 network links yield by selected interactions or relations, 382,568 participants and 253,155,020 characters in text data. Among all snapshots, 63 are ego snapshots, 54 are group snapshots; 49 have network interaction links, 89 have network friendship links; 43 have text content from messages.

# The database consists of 31,996,740 triples, 3,030,433 network links yield by main interactions or relations, 350,284 participants and 245,194,377 characters. Among all snapshots, 63 are ego snapshots, 54 are group snapshots; 50 have interaction edges, 89 have friendship edges; 43 have text content from messages.